---
title: "What did presidents react towards recessions?"
author: "Jing Zhao jz2786"
output:
  html_document: default
  html_notebook: default
---
 
 
 The word "Recessions" always makes people feel desperate and painful. However, assume recession happens or is happening when a nominee is running for president, what would he or she think? Will he/she feel fear? Or in contrast, will the recession just stimulate he/she to generate a sense of mission.
 
 I pick up 6 periods of recession. All these recessions just happened during the elections. They are Panic of 1873, 1882-85 recession, depression of 1920-21, 1929 great depression, 1980 recession and 2007 great depression. Accordingly, these presidents are Rutherford B. Hayes, Grover Cleveland, Warren G. Harding, Franklin D. Roosevelt, Ronald Reagan and Barack Obama.
 
 I will use sentiment analysis and topic models to analyze what kind of emotions presidents during recession time deliver and whether they pay more attention on economics related topics.


## 1. Preparation: install and load packages 
```{r}
packages.used=c("rvest", "tibble", "qdap", 
                "sentimentr", "gplots", "dplyr",
                "tm", "syuzhet", "factoextra", 
                "beeswarm", "scales", "RColorBrewer",
                "RANN", "tm", "topicmodels")

# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}
install.packages("tidytext",dependencies=TRUE,repos='http://cran.rstudio.com/')
# load packages
library("rvest")
library("tibble")
# You may need to run
# sudo ln -f -s $(/usr/libexec/java_home)/jre/lib/server/libjvm.dylib /usr/local/lib
# in order to load qdap
library("qdap")
library("sentimentr")
library("gplots")
library("dplyr")
library("tm")
library("syuzhet")
library("factoextra")
library("beeswarm")
library("scales")
library("RColorBrewer")
library("RANN")
library("tm")
library("topicmodels")

source("../lib/plotstacked.R")
source("../lib/speechFuncs.R")
```

```{r}
print(R.version)
```

## 2. Data for Sentiment Analysis
```{r,message=FALSE, warning=FALSE}
inaug.folder.path="../data/fulltext/"

inaug.all.speeches=list.files(path = inaug.folder.path, pattern ="^inaug.*\\.txt$") 
inaug.all.list=read.csv("../data/inauglist.csv", stringsAsFactors = FALSE)
speech.all.list = inaug.all.list
speech.all.list$type=rep("inaug", nrow(inaug.all.list))
for(i in seq(nrow(speech.all.list))){
  speech.all.list$fulltext[i] =
    readLines(paste("../data/fulltext/",inaug.all.speeches[i],sep=''),
                                          n=-1, skipNul=TRUE)
}
```


```{r,message=FALSE, warning=FALSE}
presidents.recession=c("BarackObama","FranklinDRoosevelt","GroverCleveland","RonaldReagan","RutherfordBHayes","WarrenGHarding")
sentence.list=NULL
for(i in 1:nrow(speech.all.list)){
  sentences=sent_detect(speech.all.list$fulltext[i],
                        endmarks = c("?", ".", "!", "|",";"))
  if(length(sentences)>0){
    emotions=get_nrc_sentiment(sentences)
    word.count=word_count(sentences)
    # colnames(emotions)=paste0("emo.", colnames(emotions))
    # in case the word counts are zeros?
    emotions=diag(1/(word.count+0.01))%*%as.matrix(emotions)
    sentence.list=rbind(sentence.list, 
                        cbind(speech.all.list[i,-ncol(speech.all.list)],
                              sentences=as.character(sentences), 
                              word.count,
                              emotions,
                              sent.id=1:length(sentences)
                              )
    )
  }
}
sentence.list=
  sentence.list%>%
  filter(!is.na(word.count)) 
```

##3. Comparsion with recession background
Here I am going to compare sentiments in inaugurals between presidents during recessions and the remaining presidents.
```{r}
presidents.normal = setdiff(speech.all.list$File,presidents.recession)

heatmap.2(cor(sentence.list%>%filter(type=="inaug", File%in%presidents.recession)%>%select(anger:trust)), 
          scale = "none", 
          col = bluered(100), , margin=c(6, 6), key=F,
          trace = "none", density.info = "none")
```

```{r}
heatmap.2(cor(sentence.list%>%filter(type=="inaug",File%in%presidents.normal)%>%select(anger:trust)), 
          scale = "none", 
          col = bluered(100), , margin=c(6, 6), key=F,
          trace = "none", density.info = "none")

```

```{r}


emo.means.recession=colMeans(select(sentence.list%>%filter(type=="inaug",File%in%presidents.recession), anger:trust)>0.01)
col.use=c("red2", "darkgoldenrod1", 
            "chartreuse3", "blueviolet",
            "darkgoldenrod2", "dodgerblue3", 
            "darkgoldenrod1", "darkgoldenrod1")
barplot(emo.means.recession[order(emo.means.recession)], las=2, col=col.use[order(emo.means.recession)], horiz=T, main="Inaugural Speeches under recession")


```

```{r}
emo.means=colMeans(select(sentence.list%>%filter(type=="inaug",File%in%presidents.normal), anger:trust)>0.01)
barplot(emo.means[order(emo.means)], las=2, col=col.use[order(emo.means)], horiz=T, main="Inaugural Speeches under normal economic condition")
```


We can find a significant difference between these two figures is fear. Obviously, presidents
who won under serious recession background had more fear emotion and less surprise and anticipation here. In other words, it also means that under special condition, presidents might feel uncertain about future. They might be concerned about how to recover the economy and whether the policies they proposed would work. 
 
 
```{r}
presid.summary=tbl_df(sentence.list)%>%
  filter(type=="inaug", File%in%presidents.recession)%>%
  #group_by(paste0(type, File))%>%
  group_by(File)%>%
  summarise(
    anger=mean(anger),
    anticipation=mean(anticipation),
    disgust=mean(disgust),
    fear=mean(fear),
    joy=mean(joy),
    sadness=mean(sadness),
    surprise=mean(surprise),
    trust=mean(trust)
    #negative=mean(negative),
    #positive=mean(positive)
  )

presid.summary=as.data.frame(presid.summary)
rownames(presid.summary)=as.character((presid.summary[,1]))
km.res=kmeans(presid.summary[,-1], iter.max=200,
              2)
fviz_cluster(km.res, 
             stand=F, repel= TRUE,
             data = presid.summary[,-1], xlab="", xaxt="n",
             show.clust.cent=FALSE)

```

I used K-means method to find that Warren G. Harding had a different sentiment from other 5 presidents. Through preside.summary, we can find president Harding showed more positive emotions in his inaugurals. For instance, there are much more joy and trust in his inaugurals and less disgust.


## 4. Topic Models: Get the speech data
```{r}
folder.path="../data/president/"

inaug.speeches=list.files(path = folder.path, pattern ="^inaug.*\\.txt$") 
inaug.prex.out=substr(inaug.speeches, 6, nchar(inaug.speeches)-4)

inaug.ff.all<-Corpus(DirSource(folder.path))

```


```{r}
inaug.ff.all<-tm_map(inaug.ff.all, stripWhitespace)
inaug.ff.all<-tm_map(inaug.ff.all, content_transformer(tolower))
inaug.ff.all<-tm_map(inaug.ff.all, removeWords, stopwords("english"))
inaug.ff.all<-tm_map(inaug.ff.all, removeWords, character(0))
inaug.ff.all<-tm_map(inaug.ff.all, removePunctuation)
inaug.dtm.all<-t(TermDocumentMatrix(inaug.ff.all))
#inaug.dtm.tidy<-tidy(inaug.dtm.all)
#inaug.dtm.overall=summarise(group_by(inaug.dtm.tidy, term), sum(count))
```

## 5. wordcloud with IF-IDF weighting
```{r,message=FALSE, warning=FALSE}
library(tidyr)
library(tidytext)
library(wordcloud)
library(RColorBrewer)
inaug.dtm.all.if.idf <- DocumentTermMatrix(inaug.ff.all,
                          control = list(weighting = function(x)
                                             weightTfIdf(x, 
                                                         normalize =FALSE),
                                         stopwords = TRUE))
ff.dtm.if.idf=tidy(inaug.dtm.all.if.idf)
idf.tdm.overall=summarise(group_by(ff.dtm.if.idf, term), sum(count))
wordcloud(idf.tdm.overall$term, idf.tdm.overall$`sum(count)`,
          scale=c(4,0.4),
          max.words=100,
          min.freq=1,
          random.order=FALSE,
          rot.per=0.3,
          use.r.layout=T,
          random.color=FALSE,
          colors=brewer.pal(9,"Blues"))

```

IF-IDF weighting can help us to find more interesting words presidents wanted to emphasize. Thus here our 6 presidents wanted to focus on 'america' and 'helped'. I think it's probably because during the hard times, citizens need to get united to conquer difficulties together. They need to make America stronger again.

##6. Five Topics' meanings by using LDA (each topic has a mixture of words)
```{r}
require(quanteda)
require(slam)

topicmodel1<-LDA(inaug.dtm.all, 5)
inaug.topics <- tidy(topicmodel1)
inaug_top_terms <- inaug.topics %>%
  group_by(topic) %>%
  top_n(15, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

inaug_top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()



```

In these topics, we can briefly describe different topics.
For instance, Topic 1 involves more about politics and government; topic 2 includes more about america; topic 3 might talk about economy; topic 4 involves laws and social service; topic 5 focuses on people. Since the corpus has many words like 'will' and 'the', it could be better to understand the 15 topics if we remove them before we build the LDA model.

##7. Each document has  mixture of topics
```{r}
inaug.documents <- tidy(topicmodel1, matrix = 'gamma')


inaug.documents %>%
  mutate(document = reorder(document, gamma * topic)) %>%
  ggplot(aes(factor(topic), gamma)) +
  geom_boxplot() +
  facet_wrap(~ document)
```

The figures above shows that each document has a mixture of topics. And President Reagan focused on topic 2 (America) in his inaugural; President Harding focused on topic 2 too; President Obama focused on topic 5 (people); President Hayes focused on topic 1 (politics and government); President Roosevelt emphasized topic 3 (economy) and President Cleveland pay attention to topic 4(laws). It is not surprise that President Roosevelt's inaugural mainly involves economy topic since 1929 great recession is really impressive for every one. 

##8. Comparison between Republican and Democratian
```{r}
party.path="../data/comparison/"

party.speeches=list.files(path = party.path) 

party.ff.all<-Corpus(DirSource(party.path))

party.ff.all<-tm_map(party.ff.all, stripWhitespace)
party.ff.all<-tm_map(party.ff.all, content_transformer(tolower))
party.ff.all<-tm_map(party.ff.all, removeWords, stopwords("english"))
party.ff.all<-tm_map(party.ff.all, removeWords, character(0))
party.ff.all<-tm_map(party.ff.all, removePunctuation)
party.dtm.all<-t(TermDocumentMatrix(party.ff.all))

topicmodel2<-LDA(party.dtm.all, 5)
party.topic = tidy(topicmodel2)
party_top_terms <- party.topic %>%
  group_by(topic) %>%
  top_n(20, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)
party_top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()

party.documents <- tidy(topicmodel2, matrix = 'gamma')


party.documents %>%
  mutate(document = reorder(document, gamma * topic)) %>%
  ggplot(aes(factor(topic), gamma)) +
  geom_boxplot() +
  facet_wrap(~ document)


```

We concluded the topics above are politics & government, People, Economy, America & nation and laws & international affairs. 
Here we can find that different parties indeed affect their inaugurals topic. Because different parties have their own policy to govern, and their proposals for policy are quite diversified even they might be under a similar economic situation.


##9. Conclusion

1. During economic recession days, presidents delivered more fear emotion and less surprise and anticipation emotion in their inaugurals. It means elected as a president in the hard time indeed seems like a challenge for them to take. But less surprise demonstrated that they also seemed to know they are the ones to save the country.
 
2. Presidents during recessions prefer the word 'america' and other words which can inspire and encourage citizens in order to make people feel hopeful about their future. They might try to use the spirit of the nation to motivate them to realize their American Dream.

3. By using LDA model to get 5 topics from presidents' inaugurals, we find President Roosevelt foucsed on economy while other presidents emphasize other aspects. However, it is not hard to understand because the Great Depression in 1929 was indeed horrible compared to other recessions. Recovering economy was the priority for the new government. 

4. When we want to compare whether there exist any difference between the two parties in their inaugurals, the result shows that it did. Republicans prefer to talk about administration and laws while Democrats prefer to discuss America and economy when the economic background of United States are not optimistic.